Integrated approach to modeling krill population dynamics in the Western Antarctic Peninsula: spatial and ecosystem considerations

### 2024-09-11 Observations_SAM_SS3_Krill_481.Rmd
test n
finish edits (i hope)

---
title: "APPENDIX 1: Integrated approach to modeling krill population dynamics in Western Antarctic Peninsula. Spatial and ecosystem considerations"
subtitle: "Observations WP- SAM 2024/26"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::pdf_document2:
    number_sections: false
fig_caption: yes
bibliography: SA_krill.bib
csl: apa.csl
link-citations: yes
toc: false
linkcolor: blue
linestretch: 1.3
header-includes:
- \fontsize{12}{16}
- \selectfont
- \usepackage{lscape}
---
\newpage
\tableofcontents
\newpage

@@

The main recommendations were extracted from @CCAMLRWGSAM2024.

@@

We consider that a length-based model as an alternative to the age-based approach, especially given the scarcity of direct age data in krill monitoring. Aging techniques for krill have not been available historically, although some recent progress has been made [@Kilada2017]. In this sense, when reading of hard parts is difficult like in krill, catch-at-length data are more plentiful, as the collection of length information is relatively cheap. Length offers insights into the age structure of the population, as there is a correlation between age and length.  So far, all integrated models for krill have been of the Catch-at-Length type [@Kinzey2015a; @Kinzey2019a; @Wang2021]. This means they are approaches that internally model age through growth parameters and the Age-length-key (ALK). To perform this conversion, various methods have been used, but the AKL matrix is perhaps the most important one [@ICCA2003].

@@

The proportions-at-age are then adjusted to find the best fit between the observed size frequency data and that predicted by the proportion-at-age and the ALK. The majority of ALKs are based on growth parameters according to the most suitable formula for the species. In the case of krill, these parameters are derived from the von Bertalanffy growth curve. In @Mardones2023, using `SS3` stock assessment catch-at-lenght model, we use a von Bertalanffy growth relationship with estimated parameters L~inf~, k, and CV. This length-at-age relationship, mediated by survey catchability and selectivity, was used to compare survey biomasses with model estimates of vulnerable biomass (Eq. (A.4)). Weight-at-length was assumed known and calculated using @Maschette2020 parametres.

@@

```{r platoon, out.width='70%', fig.cap="Reprsentation of platoons in Krill lenght frencuecies"}
knitr::include_graphics('Figs/Plattoons.png')
```

@@

Since we don't have age data monitoring program, we plan to explore *platoon* alternative to compare its performance and results with the current length-based model using currently in WG-SAM-2024/26.

@@

Regarding @WGSAM2114, they provide diagnostic plots for the 2021 assessment model for Antarctic toothfish (*Dissostichus mawsoni*) in the Ross Sea region, following the recommendations of WG-SAM-2015 (SC-CAMLR-XXXIV 2015 Annex 5), and for which it is recommended to follow the diagnostic structure for this analysis. @WGSAM2114 has several way to test the perfomance of stock assessment in toothfish, among them are;  progrssion catch-at-age, likelihood profile, Maximum posterior density (MPD), Autocorrelation plots (ACF), Pearson residuals by sex, Markov Chain Mote Carlo priors and posteriors distributions and fits to ages compositions.

@@

```{r acf, out.width='90%', fig.cap="Cross correlation in autoregresive analysis for all scenarios in krill population"}
knitr::include_graphics('index_files/figure-html/unnamed-chunk-46-1.jpeg')
```

@@

While we acknowledge the work done by @WGSAM2114 in toothfish stock assessment, we propose a converge towards these standard practices, regardless of the assessment platform and the structure of the integrated model.

@@

This observation raise implications for fishery selectivity. To develop this we configured  a selectivity block in an alternative scenario (`s9`) and compared its performance with the base model (`s1`). This is configured for all strata in 48.1. Tn this case, selectivity blocks are configured in `SS3` as show this template lines in `control` file:

@@

```{r comparebio, out.width='80%', fig.cap="biomass estimates  in s1 and s9 scenarios."}
knitr::include_graphics('Figs/COM1_9compare4_Bratio_uncertainty.png')
```

```{r comparerec, out.width='80%', fig.cap="biomass estimates  in s1 and s9 scenarios."}
knitr::include_graphics('Figs/COM1_9compare10_recruits_uncertainty.png')
```
## Observation 3.2

@@

There are two important issues related to the current stock assessment approach presented in WG-SAM 24/27. The first one, while we understand that there are krill flows to and from Area 48.1, the challenge of having time series data to implement a model that covers most of Subarea 48 becomes difficult. However, we note that as monitoring programs provide the necessary data to implement models of this nature, we are open to a more comprehensive approach in the Southern Ocean. Secondly, the current model implemented in `SS3` considered the spatial heterogeneity of the krill stock structure within Area 48.1, as illustrated in the Figure \@ref(fig:ksh), however, a model that involves movement between strata has not yet been developed, and is part of the ongoing work presented at the SAM.

@@

```{r trawl, out.width='70%', fig.cap="Display of leght compositios with different trawl technique use in 48.1 subarea."}
knitr::include_graphics('Figs/trawl.png')
```

@@

We agree that a progression from simple to more complex model implementations would facilitate understanding and provide a clearer evaluation of the model assumptions. However, and regarding this observation, a visual inspection of Pearson residuals can lead to a mistaken conclusion about the model *goodness-of-fit*. We have integrated a quantitative element to evaluate this type of test. The plotting options are kept mainly to those provided by `r4ss`. With `SSplotRunstest()`, runs tests results are illustrated for mean lengths of size composition data from the  `SS3`model. Green shading indicates
no evidence (p > 0.05) and red shading evidence (p < 0.05) to reject the hypothesis of a randomly distributed time-series of residuals, respectively. The shaded (green/red) area spans three residual standard deviations to either side from zero, and the red points outside of the shading violate the ‘three-sigma limit’ for that series. We plot this in Figure \@ref(fig:runtest).

@@

Regarding the reliability of the parameters, it is important to note that the model is conditioned with input parameters gathered from the literature [@Smith2023a; @Maschette2020; @Kinzey2011], as is showed in Table \@ref(tab:parainit). However, the model calculates the most essential parameters through an iterative process, ensuring that their correlation is carefully managed. This calculation process is specifically designed to avoid dependence on the initial conditioning parameters of the model.

@@

Although it was not recorded in the WGSAM-2024 report, one of the questions raised was related to the representation of population variables, such as the harvest rate. In this document, we present them in Figure \@ref(fig:hr). for the entire analyzed series and for each stratum.

>>>>

---
title: "APPENDIX 1: Integrated approach to modeling krill population dynamics in Western Antarctic Peninsula. Spatial and ecosystem considerations"
subtitle: "Response to observations provided by the Working Group on WP- SAM 2024/26"
output:
  bookdown::pdf_document2:
    number_sections: false
fig_caption: yes
bibliography: SA_krill.bib
csl: apa.csl
link-citations: yes
toc: false
linkcolor: blue
linestretch: 1.3
header-includes:
- \fontsize{12}{16}
- \selectfont
- \usepackage{lscape}
---
\newpage
\tableofcontents
\newpage

@@

The main recommendations were extracted from the WG-SAM report @CCAMLRWGSAM2024.

@@

We consider a length-based model as an alternative to the age-based approach, especially given the scarcity of direct age data in krill monitoring. Aging techniques for krill have not been available historically, although some recent progress has been made [@Kilada2017]. In this sense, when reading of hard parts is difficult like in krill, catch-at-length data are more plentiful, as the collection of length information is relatively cheap. Length offers insights into the age structure of the population, as there is a correlation between age and length.  So far, all integrated models for krill have been of the Catch-at-Length type [@Kinzey2015a; @Kinzey2019a; @Wang2021]. This means they are approaches that internally model age through growth parameters and the Age-length-key (ALK). To perform this conversion, various methods have been used, but the AKL matrix is perhaps the most important one [@ICCA2003].

@@

The proportions-at-age are then adjusted to find the best fit between the observed size frequency data and that predicted by the proportion-at-age and the ALK. The majority of ALKs are based on growth parameters according to the most suitable formula for the species. In the case of krill, these parameters are derived from the von Bertalanffy growth curve. In @Mardones2023, using `SS3` stock assessment catch-at-lenght model, we use a von Bertalanffy growth relationship with estimated parameters L~inf~, k, and CV. This length-at-age relationship, mediated by survey catchability and selectivity, was used to compare survey biomasses with model estimates of vulnerable biomass (Eq. (A.4)). Weight-at-length was assumed known and calculated using @Maschette2020 parameters.

@@

```{r platoon, out.width='70%', fig.cap="Representation of platoons in Krill length frequencies"}
knitr::include_graphics('Figs/Plattoons.png')
```

@@

Since we do not have an age data monitoring program, we explored a *platoon* alternative to compare its performance and results with the current length-based model currently in WG-SAM-2024/26 [@Mardones2023].

@@

Regarding @WGSAM2114, they provide diagnostic plots for the 2021 assessment model for Antarctic toothfish (*Dissostichus mawsoni*) in the Ross Sea region, following the recommendations of WG-SAM-2015 (SC-CAMLR-XXXIV 2015 Annex 5), and for which it is recommended to follow the diagnostic structure for this analysis. @WGSAM2114 has several way to test the perfomance of stock assessment in toothfish, among them are;  progrssion catch-at-age, likelihood profile, Maximum posterior density (MPD), Autocorrelation plots (ACF), Pearson residuals by sex, Markov Chain Monte Carlo priors and posteriors distributions and fits to ages compositions.

@@

```{r acf, out.width='90%', fig.cap=" Cross correlation in autoregressive analysis for all scenarios in krill population"}
knitr::include_graphics('index_files/figure-html/unnamed-chunk-46-1.jpeg')
```

@@

While we acknowledge the work done by @WGSAM2114 in toothfish stock assessment, we propose a convergence towards these standard practices, regardless of the assessment platform and the structure of the integrated model.

@@

This observation raises implications for fishery selectivity. To develop this we configured  a selectivity block in an alternative scenario (`s9`) and compared its performance with the base model (`s1`). This is configured for all strata in 48.1. Tn this case, selectivity blocks are configured in `SS3` as show this template lines in `control` file:

@@

```{r comparebio, out.width='80%', fig.cap="Biomass estimates  in s1 and s9 scenarios."}
knitr::include_graphics('Figs/COM1_9compare4_Bratio_uncertainty.png')
```

```{r comparerec, out.width='80%', fig.cap="Biomass estimates  in s1 and s9 scenarios."}
knitr::include_graphics('Figs/COM1_9compare10_recruits_uncertainty.png')
```
## Observation 3.2

@@

There are two important issues related to the current stock assessment approach presented in WG-SAM 24/27. The first one, while we understand that there are krill flows to and from Subrea 48.1, the challenge of having time series data to implement a model that covers most of Area 48 becomes difficult. However, we note that as monitoring programs provide the necessary data to implement models of this nature, we are open to a more comprehensive approach in the Southern Ocean. Secondly, the current model implemented in `SS3` considered the spatial heterogeneity of the krill stock structure within Subrea 48.1, as illustrated in the Figure \@ref(fig:ksh), however, a model that involves movement between strata has not yet been developed, and is part of the ongoing work presented at WG-SAM.

@@

```{r trawl, out.width='70%', fig.cap="Display of length compositions with different trawl technique use in 48.1 subarea."}
knitr::include_graphics('Figs/trawl.png')
```

@@

We agree that a progression from simple to more complex model implementations would facilitate understanding and provide a clearer evaluation of the model assumptions. However, and regarding this observation, a visual inspection of Pearson residuals can lead to a mistaken conclusion about the model *goodness-of-fit*. We have integrated a quantitative element to evaluate this type of test. The plotting options are kept mainly to those provided by `r4ss`. With `SSplotRunstest()`, run tests results are illustrated for mean lengths of size composition data from the  `SS3`model. Green shading indicates
no evidence (p > 0.05) and red shading evidence (p < 0.05) to reject the hypothesis of a randomly distributed time-series of residuals, respectively. The shaded (green/red) area spans three residual standard deviations to either side from zero, and the red points outside of the shading violate the ‘three-sigma limit’ for that series. We plot this in Figure \@ref(fig:runtest).

@@

Regarding the reliability of the parameters, it is important to note that the model is conditioned with input parameters gathered from the literature [@Smith2023a; @Maschette2020; @Kinzey2011], as is shown in Table \@ref(tab:parainit). However, the model calculates the most essential parameters through an iterative process, ensuring that their correlation is carefully managed. This calculation process is specifically designed to avoid dependence on the initial conditioning parameters of the model.

@@

Although it was not recorded in the WGSAM-2024 report, one of the questions raised was related to the representation of population variables, such as the harvest rate.  The historical harvest rate of krill, as depicted in the graph, shows a significant decline from the 1970s through the 1990s, followed by a period of stabilization at much lower levels in the 2000s and 2010s. Initially, the exploitation rate was notably high, particularly in the Bransfield stratum, indicating intense harvesting activities during these early decades. However, after the early 1990s, the exploitation rate decreased sharply, suggesting the implementation of more effective management practices or a reduction in fishing pressure. In recent years, the exploitation rate across all strata has remained relatively low and stable, which may reflect sustainable fisheries management efforts. This trend highlights the importance of continued monitoring and adaptive management to ensure the long-term sustainability of krill populations in the face of environmental changes and evolving fishing practices (Figure \@ref(fig:hr)).

### 2024-09-11 SS3_Krill_481.Rmd
testc4
finish edits (i hope)

---
title: "Integrated approach to modeling krill population dynamics in Western Antarctic Peninsula. Spatial and ecosystem considerations"
author:
  - Mauricio Mardones^[Universidad de Magallanes, mamardon@umag.cl, Milenio Base]
  - Lucas Krüger ^[Instituto Antártico Chileno, Milenio Base]
  - Francisco Santa Cruz ^[Instituto Antártico Chileno, Universidad de Magallanes, Milenio Base]
  - César Cárdenas^[Instituto Antártico Chileno, Milenio Base]
  - George Watters^[Ecosystem Science Division, Southwest Fisheries Science Center, NOAA, USA]

date:  "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::pdf_document2:
    number_sections: false
fig_caption: yes
bibliography: SA_krill.bib
csl: apa.csl
link-citations: yes
toc: false
linkcolor: blue
linestretch: 1.3
header-includes:
- \fontsize{12}{16}
- \selectfont
- \usepackage{lscape}
---
\newpage
\tableofcontents
\newpage

>>>>

---
title: "Integrated approach to modeling krill population dynamics in the Western Antarctic Peninsula: spatial and ecosystem considerations"
author:
  - Mauricio Mardones^[Universidad de Magallanes, mamardon@umag.cl, Instituto Milenio Base]
  - Lucas Krüger ^[Instituto Antártico Chileno, Instituto Milenio Base]
  - Francisco Santa Cruz ^[Instituto Antártico Chileno, Universidad de Magallanes, Instituto Milenio Base]
  - César Cárdenas^[Instituto Antártico Chileno, Instituto Milenio Base]
  - George Watters^[Ecosystem Science Division, Southwest Fisheries Science Center, NOAA, USA]
output:
  bookdown::pdf_document2:
    number_sections: false
fig_caption: yes
bibliography: SA_krill.bib
csl: apa.csl
link-citations: yes
toc: false
linkcolor: blue
linestretch: 1.3
header-includes:
- \fontsize{12}{16}
- \selectfont
- \usepackage{lscape}
---
\newpage
\tableofcontents
\newpage
